Symmetric Peer-to-Peer Applications

### 2022-10-28 tpd-22.tex
...

In Section~\ref{sec.related}, we revisit existing solutions for symmetric
distributed applications.
In Section~\ref{sec.tml}, we detail our proposed middleware.
In Section~\ref{sec.app}, ...
In Section~\ref{sec.eval}, ...
In Section~\ref{sec.conclusion}, ...

- assumptions
    - correct (non-malicious) nodes

- local first
- p2p
- less restrictive (compared to CRDTs)
- semantic events
    - limitation b/c of rollbacks
    - features b/c of network traffic

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
%
We run the application main loop at $50 FPS$ ($20ms$ per frame) and calculate
the middleware overhead at each iteration, which is $9us$ in the average,
corresponding to less than $1/2000$ of the CPU time.
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency ($5-500ms$ between each peer),
    (ii)  the rate of events ($5-200~evt/sec$), and
    (iii) the delay multiplier ($1-5x$).
%
Figure~\ref{fig.mult} shows the event delays considering the network latency
and chosen multiplier.
For instance, a $50ms$ latency with a $2x$ multiplier results in $500ms$ event
delays, i.e., each event is triggered with this delay to compensate the given
latency and fixed average of 5 hops.
%
These variations result in $108$ combinations.
We executed each combination $3$ times for $5$ minutes each.
We also make some modifications to evaluate peer churn, which requires to
re-execute the simulation.
The experiments take around $40$ hours to complete.

@@

We now detail the results and how we instrument the middleware implementation
to measure the properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead in time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
Figure~\ref{fig.baks} presents the results considering all combinations of
parameters (i), (ii), and (iii) above.
The colors indicate the feasibility of each tested scenario
    (\emph{white=good}, \emph{yellow=moderate}, \emph{red=poor},
    \emph{black=faulty}).
%
For instance, the three framed rectangles ($0.1$, $3.1$, and $7.2$) in the
figure represent the scenarios with $50ms$ network delay, generating
$25~evt/min$.
Each rectangle uses a different delay multiplier, showing a good performance
for $5x$, moderate for $2x$, and poor for $1x$.
%
\item \textbf{Real-time pace of peers:}
All peers output their current ticks continuously, allowing us to compare them
in the shared timeline.
Every time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table akin to Figure~\ref{fig.baks}), which
would not provide further insights.
%
\item \textbf{Correctness of unstable peers:}
We tweak our simulation to make each peer remain $40s$ online and $20s$
offline in the average, including peers \code{9-11} in the middle of the
network.
This creates partitions in the network, which generate events out of sync that
require constant resynchronization between peers.
%
Then, to evaluate the network correctness under such high churn, we count how
much time a recovering peer takes to catch up in real time:
When a peer becomes online, we take the current maximum tick considering all
reachable peers, and then count the time the recovering peer takes to catch up.
We also assert that they all reach the same final state eventually.
%
Considering all scenarios, the average is $1.3$ seconds of recovering time,
with no significant variance.
Therefore, we also omit the resulting table (akin to Figure~\ref{fig.baks}).
%
\end{enumerate}

@@

Our results show that goals (b) and (c) are met in all scenarios consistently,
and we consider them nearly optimal:
Regardless of the network latency, event rate, and delay multiplier,
    the variation in the pace of peers is under $0.2\%$ (goal b), while
    recovering peers take $1.3$ seconds to catch up in time (goal c).
Also regarding goal (c), note that the middleware makes smooth time transitions
in small steps that take $1$ second to complete.
Therefore, the average of $1.3$ seconds shows that peers need no more than 2
time travels while recovering.

@@

>>>>

In Section~\ref{sec.related}, we revisit existing solutions for symmetric
distributed applications.
In Section~\ref{sec.tml}, we detail the architecture of our middleware, its
programming API, and how it orchestrates peers.
%In Section~\ref{sec.app}, ...
In Section~\ref{sec.eval}, we evaluate our design under a number of scenarios.
In Section~\ref{sec.conclusion}, we conclude this work.

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
%
We run the application main loop at $50~FPS$ ($20ms$ per frame) and calculate
the middleware overhead at each iteration, which is $9us$ in the average,
corresponding to less than $1/2000$ of the CPU time.
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency ($5-500ms$ between each peer),
    (ii)  the rate of events ($5-200~evt/sec$), and
    (iii) the delay multiplier ($1-5x$).
%
Figure~\ref{fig.mult} shows the event delays considering the network latency
and chosen multiplier.
For instance, a $50ms$ latency with a $2x$ multiplier results in $500ms$ event
delays, i.e., each event is triggered with this delay to compensate the given
latency and fixed average of 5 hops.
%
These variations result in $108$ combinations.
We executed each combination $3$ times for $5$ minutes each.
We also make some modifications to evaluate peer churn, which requires to
re-execute the simulation.
The experiments take around $40$ hours to complete.

% https://docs.google.com/spreadsheets/d/1xK5_wF9MoR-SmPGWVj0sGLrHxhpPBlLjVoS8U-Y7zTQ/edit#gid=63016757

@@

We now detail the results and how we instrument the middleware implementation
to measure the properties above:

\textbf{(a) Recurrence of time travels:}
Every time a peer is ahead in time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
Figure~\ref{fig.baks} presents the results considering all combinations of
parameters (i), (ii), and (iii) above.
The colors indicate the feasibility of each tested scenario
    (\emph{white=good}, \emph{yellow=moderate}, \emph{red=poor},
    \emph{black=faulty}).
%
For instance, the three framed rectangles ($0.1$, $3.1$, and $7.2$) in the
figure represent the scenarios with $50ms$ network delay, generating
$25~evt/min$.
Each rectangle uses a different delay multiplier, showing a good performance
for $5x$, moderate for $2x$, and poor for $1x$.

\textbf{(b) Real-time pace of peers:}
All peers output their current ticks continuously, allowing us to compare them
in the shared timeline.
Every time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table akin to Figure~\ref{fig.baks}), which
would not provide further insights.

\textbf{(c) Correctness of unstable peers:}
We tweak our simulation to make each peer remain $40s$ online and $20s$
offline in the average, including peers \code{9-11} in the middle of the
network.
This creates partitions in the network, which generate events out of sync that
require constant resynchronization between peers.
%
Then, to evaluate the network correctness under such high churn, we count how
much time a recovering peer takes to catch up in real time:
When a peer becomes online, we take the current maximum tick considering all
reachable peers, and then count the time the recovering peer takes to catch up.
We also assert that they all reach the same final state eventually.
%
Considering all scenarios, the average is $1.3$ seconds of recovering time,
with no significant variance.
Therefore, we also omit the resulting table (akin to Figure~\ref{fig.baks}).

@@

Our results show that goals (b) and (c) are met in all scenarios consistently,
and we consider them nearly optimal:
Regardless of the network latency, event rate, and delay multiplier,
    the variation in the pace of peers is under $0.2\%$ (goal b), while
    recovering peers take $1.3$ seconds to catch up in time (goal c).
Note that the middleware makes smooth time transitions in small steps that take
$1$ second to complete, resulting in no more than 2 time travels to recover
peers.

@@

We propose a middleware to program \emph{symmetric peer-to-peer applications}
in which interacting decentralized instances conform to identical behavior.
%
The middleware API shares similar limitations with centralized
alternatives~\cite{gals,croquet}, being restricted to deterministic and
stateless calls, and only supports pre-allocated memory.
%
Our main contribution is a ``time machine'' based on memory snapshots and
deterministic simulation, allowing to decentralize the network timeline such
that conflicting peers can rollback and resynchronize their state.
%
We assume that peers are non-malicious and that they can assign themselves
unique identifiers in a contiguous range.

The middleware can handle applications with over 20 peers in a heterogeneous
topology under high churn.
%
We show that peers can
    (a) broadcast events and meet deadlines globally, 
    (b) advance in sync in real time, and
    (c) leave and join the network and remain symmetric.
%
In our experiments, we vary the network latency and the rate and deadline of
events.
We than measure the recurrence of time travels in the network as a
representative of quality.
%
As a result, we present a feasibility table to indicate the application
scenarios that the middleware can handle.
For instance, in a simulation scenario of 25 events per minute with a network
latency of $50ms$, the peers need to roll back and resynchronize only $3\%$ of
the time.
